### 2018-10-04 notebooks/article_notebook.Rmd
minor changes

---
title: "Article data analysis"
author: "Håkon Kaspersen"
date: '`r format(Sys.time(), "%d.%m.%Y")`'
output:
    html_document:
      code_folding: hide
      df_print: paged
      highlight: kate
      number_sections: yes
      theme: flatly
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
    pdf_document:
      toc: yes
      toc_depth: '3'
---

# Dependencies
This notebook is using the following packages:

- dplyr

- ggplot2

- purrr
```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable)
```

```{r, include=FALSE}
id <- c("2017-01-3941-1-S", "2016-17-98-1-S", "2016-17-610-1-S", "2016-17-604-1-S", 
"2016-17-565-1-S", "2016-17-550-1-S", "2016-17-548-1-S", "2016-17-546-1-S", 
"2016-17-485-1-S", "2016-17-469-1-S", "2016-17-423-1-S", "2016-17-412-1-S", 
"2016-17-408-1-T", "2016-17-408-1-S", "2016-17-393-1-S", "2016-17-391-1-S", 
"2016-17-388-1-T", "2016-17-388-1-S", "2016-17-377-1-S", "2016-17-363-1-S", 
"2016-17-352-1-S", "2016-17-343-1-S", "2016-17-340-1-S", "2016-17-336-1-S", 
"2016-17-329-1-S", "2016-17-299-1-S", "2016-17-298-1-S", "2016-17-294-1-S", 
"2016-17-292-1-S", "2016-17-288-1-S", "2016-17-281-1-T", "2016-17-281-1-S", 
"2016-17-275-1-S", "2016-17-271-1-S", "2016-17-27-1-S", "2016-17-264-1-S", 
"2016-17-263-1-S", "2016-17-257-1-S", "2016-17-256-1-S", "2016-17-242-1-S", 
"2016-17-225-1-S", "2016-17-224-1-S", "2016-17-189-1-T", "2016-17-183-1-S", 
"2016-17-174-1-T", "2016-17-174-1-S", "2016-17-164-1-S", "2016-17-16-1-S", 
"2016-17-148-1-S", "2016-17-124-1-S", "2016-17-114-1-S", "2016-17-105-1-T", 
"2016-17-1-1-S", "2016-02-785-2-S", "2016-02-783-2-T", "2016-02-771-2-S", 
"2016-02-770-2-S", "2016-02-741-2-S", "2016-02-739-2-S", "2016-02-732-2-S", 
"2016-02-727-2-S", "2016-02-726-2-S", "2016-02-725-2-S", "2016-02-723-2-S", 
"2016-02-697-2-S", "2016-02-693-2-T", "2016-02-693-2-S", "2016-02-692-2-S", 
"2016-02-690-2-T", "2016-02-690-2-S", "2016-02-620-2-S", "2016-02-563-2-S", 
"2016-02-546-2-S", "2016-02-543-2-S", "2016-02-531-2-S", "2016-02-522-2-S", 
"2016-02-515-2-T", "2016-02-515-2-S", "2016-02-514-2-S", "2016-02-513-2-T", 
"2016-02-513-2-S", "2016-02-505-2-S", "2016-02-492-2-S", "2016-02-490-2-S", 
"2016-02-486-2-S", "2016-02-447-2-S", "2016-02-446-2-S", "2016-02-428-2-S", 
"2016-02-425-2-S", "2016-02-419-2-S", "2016-02-418-2-S", "2016-02-415-2-S", 
"2016-02-413-2-T", "2016-02-413-2-S", "2016-02-411-2-S", "2016-02-410-2-S", 
"2016-02-409-2-S", "2016-02-408-2-S", "2016-02-402-2-S", "2016-02-381-2-S", 
"2016-02-369-2-S", "2016-02-366-2-S", "2016-02-363-2-S", "2016-02-359-2-S", 
"2016-02-358-2-S", "2016-02-356-2-S", "2016-02-353-2-S", "2016-02-352-2-S", 
"2016-02-351-2-S", "2016-02-348-2-S", "2016-02-345-2-T", "2016-02-345-2-S", 
"2016-02-343-2-S", "2016-02-336-2-S", "2016-02-330-2-S", "2016-02-329-2-S", 
"2016-02-324-2-S", "2016-02-316-2-S", "2016-02-302-2-S", "2016-02-293-2-S", 
"2016-02-281-2-S", "2016-02-278-2-S", "2016-01-3763-1-T", "2016-01-3474-1-T", 
"2015-01-753-1-S", "2015-01-752-1-S", "2015-01-751-1-S", "2015-01-750-1-S", 
"2015-01-749-1-S", "2015-01-5263-1-S", "2015-01-5262-1-S", "2015-01-5177-1-S", 
"2015-01-5141-1-S", "2015-01-5140-1-S", "2015-01-5138-1-S", "2015-01-5022-1-S", 
"2015-01-5020-1-S", "2015-01-5018-1-S", "2015-01-4988-1-S", "2015-01-4934-1-S", 
"2015-01-4932-1-S", "2015-01-4931-1-S", "2015-01-4794-1-S", "2015-01-4788-1-S", 
"2015-01-4742-1-S", "2015-01-467-1-S", "2015-01-466-1-S", "2015-01-463-1-S", 
"2015-01-4559-1-S", "2015-01-4558-1-S", "2015-01-4465-1-S", "2015-01-4463-1-S", 
"2015-01-4462-1-S", "2015-01-4421-1-S", "2015-01-4249-1-S", "2015-01-4037-1-S", 
"2015-01-3915-1-S", "2015-01-3914-1-S", "2015-01-3911-1-S", "2015-01-3887-1-S", 
"2015-01-3640-1-S", "2015-01-3639-1-S", "2015-01-3638-1-S", "2015-01-3548-1-S", 
"2015-01-3545-1-S", "2015-01-3318-1-S", "2015-01-3317-1-S", "2015-01-3102-1-S", 
"2015-01-3004-1-S", "2015-01-3001-1-S", "2015-01-2879-1-S", "2015-01-2878-1-S", 
"2015-01-2839-1-S", "2015-01-2838-1-S", "2015-01-2837-1-S", "2015-01-2836-1-S", 
"2015-01-2835-1-S", "2015-01-2771-1-S", "2015-01-2430-1-S", "2015-01-2429-1-S", 
"2015-01-2426-1-S", "2015-01-2404-1-S", "2015-01-2403-1-S", "2015-01-2309-1-S", 
"2015-01-2166-1-S", "2015-01-2097-1-S", "2015-01-2094-1-S", "2015-01-1955-1-S", 
"2015-01-1954-1-S", "2015-01-1952-1-T", "2015-01-1951-1-S", "2015-01-1950-1-S", 
"2015-01-1949-1-S", "2015-01-1826-1-S", "2015-01-1823-1-S", "2015-01-1661-1-S", 
"2015-01-1566-1-S", "2015-01-1343-1-S", "2014-01-861-1-S", "2014-01-860-1-S", 
"2014-01-856-1-S", "2014-01-7584-1-S", "2014-01-7543-1-S", "2014-01-7512-1-S", 
"2014-01-7490-1-S", "2014-01-7479-1-S", "2014-01-7443-1-S", "2014-01-7442-1-S", 
"2014-01-7394-1-S", "2014-01-7375-1-S", "2014-01-7234-1-S", "2014-01-7207-1-S", 
"2014-01-7136-1-S", "2014-01-7133-1-S", "2014-01-7050-1-S", "2014-01-6926-1-S", 
"2014-01-6924-1-S", "2014-01-6710-1-S", "2014-01-6625-1-S", "2014-01-6530-1-S", 
"2014-01-6291-1-S", "2014-01-6123-1-S", "2014-01-6122-1-S", "2014-01-6044-1-S", 
"2014-01-6043-1-S", "2014-01-6041-1-S", "2014-01-6040-1-S", "2014-01-5917-1-T", 
"2014-01-5917-1-S", "2014-01-5916-1-S", "2014-01-5915-1-S", "2014-01-5912-1-S", 
"2014-01-5792-1-S", "2014-01-5565-1-S", "2014-01-4922-1-T", "2014-01-4921-1-S", 
"2014-01-4651-1-S", "2014-01-4539-1-S", "2014-01-4271-1-S", "2014-01-4270-1-S", 
"2014-01-3802-1-S", "2014-01-3677-1-S", "2014-01-3676-1-S", "2014-01-3212-1-S", 
"2014-01-2996-1-S", "2014-01-2995-1-S", "2014-01-2992-1-S", "2014-01-2773-1-S", 
"2014-01-2728-1-S", "2014-01-2516-1-S", "2014-01-2366-1-S", "2014-01-2362-1-T", 
"2014-01-2145-1-S", "2014-01-2144-1-S", "2014-01-2071-1-S", "2014-01-2070-1-S", 
"2014-01-2069-1-S", "2014-01-1742-1-S", "2014-01-1741-1-S", "2014-01-1649-1-S", 
"2014-01-1268-1-T", "2014-01-1182-1-S", "2013-01-2024-1-T", "2012-01-838-6-T", 
"2011-01-9386-3-T", "2011-01-6927-5-T", "2011-01-3460-5-T", "2011-01-1544-2-T", 
"2010-01-991-7-T", "2009-01-551-1-T", "2009-01-4672-5-T", "2009-01-4618-2-T", 
"2009-01-4415-4-T", "2009-01-413-1-T", "2009-01-4064-1-T", "2009-01-3815-1-T", 
"2009-01-3064-2-T", "2009-01-2696-4-T", "2009-01-2521-3-T", "2009-01-175-4-T", 
"2009-01-175-2-T", "2009-01-1690-1-T", "2007-01-361-1-T", "2006-01-2821-1-T", 
"2006-01-2184-1-T")
```

```{r Import data, include = FALSE}
isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

no_of_reads <- read.table("../data/read_data.txt",
                          sep = "\t",
                          header = T,
                          stringsAsFactors = F)

```

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","aac","oqx")
mut_genes <- c("gyr","par","marR","soxS","tolC","acrR")

# adjust parameters for filtering
if (length(ac_genes) == 1) {
  if (grepl("all", ac_genes, ignore.case = TRUE) == TRUE) {
    ac_genes <- "ALL"
    } else {
      ac_genes <- ac_genes
    }
}

if (length(mut_genes) == 1) {  
  if (grepl("all", mut_genes, ignore.case = TRUE) == TRUE) {
    mut_genes <- "ALL"
    } else {
      mut_genes <- mut_genes
    }
}
# -------------------------------------- Libraries

if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2,dplyr,tidyr,gridExtra,grid,
               forcats,purrr,stringr,kableExtra,
               knitr,IRdisplay,reprex,svglite)

# -------------------------------------- Functions
# Collapses data frame to unique lines while ignoring NA's
func_paste <- function(x) paste(unique(x[!is.na(x)]), collapse = ", ")

# Confidence interval function
get_binCI <- function(x, n) as.numeric(setNames(binom.test(x,n)$conf.int*100,
                                                c("lwr", "upr")))

# Identifies filenames in input folder
file_names <- function(filepath) {
  files <- list.files(path = filepath, pattern = "amr_report.tsv")
  return(files)
}

# Import ariba data from report.tsv from chosen database used in ariba
get_ariba_data <- function(filepath) {
  files <- file_names(filepath)

  data_list <- lapply(files,
                      FUN = function(file) {
                        read.delim(
                          paste0(filepath, "/", file),
                          stringsAsFactors = F,
                          header = TRUE,
                          sep = "\t"
                        )
                      })

  names(data_list) <- files
  data <- bind_rows(lapply(data_list, function(x) map(x, as.character)), .id = "ref")
  return(data)
}

# Corrects the gene names found in the "cluster" column
fix_gene_names <- function(df) {
  genes <- unique(df$ref_name)
  new_names <- gsub("^(.*?)\\..*", "\\1", genes)
  new_names <- gsub("_", "", new_names, fixed = T)
  new_names <- gsub("-", "", new_names, fixed = T)

  gene_names <- c()
  for (i in new_names) {
    p <- paste(tolower(substring(i, 1,3)), substring(i, 4), sep = "", collapse = " ")
    gene_names <- c(gene_names,p)
  }
  df2 <- data.frame(genes,gene_names) %>%
    mutate(genes = as.character(genes)) %>%
    rename(ref_name = genes)

  df <- df %>%
    left_join(df2, by = "ref_name") %>%
    mutate(gene_names = as.character(gene_names),
           ref = gsub("(.*?)_amr_report.tsv", "\\1", ref))

  return(df)
}

# Function for selecting genes of interest and filtering the 
# columns in the data frame on the resulting vector
select_genes <- function(df, gene_string) {
  names_df <- names(df)
  x <- "ref"
  for (i in names_df) {
    for (j in gene_string) {
      if (str_detect(i, regex(j, ignore_case = T)) == TRUE) {
        x <- c(x, i)
      }
    }
  }
  df <- df %>%
    select(one_of(x))
  return(df)
}

# Allowed flags from the ARIBA report
flag_selection <- c("19","27","147","155","403",
                    "411","915","923","787","795",
                    "531","539","659","667","787","795")

# Function that returns info on flag selection
check_flags <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(flag_result = flag %in% flag_selection,
           flag_result = as.integer(flag_result),
           ref = gsub("(.*?)\\_.+", "\\1", ref),
           ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
           ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)) %>%
    rename("gene" = gene_names)
  return(df)
}

# Function that handles megares data and returns a data frame with information on whether a gene
# is mutated or not. Includes control for QRDR in gyrA.
create_mut_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    filter(flag %in% flag_selection) %>%
    mutate(id = 1:n()) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, flag)) %>%
    gather(gene, mut, -ref) %>%
    mutate(mut = ifelse(mut == "" | mut == "." | is.na(mut) == TRUE, 0, mut),
           result = ifelse(mut != 0, 1, 0),
           result = as.integer(result),
           type = "mut")
  return(df)
}

# Corrects or mutations in QRDR for gyrA, gyrB, parC and parE genes.
# The function returns columns of 1/0 values for whether or not the 
# mutations reported by ARIBA is within the QRDR in the gene:
# gyrA: AA 67 - 106
# gyrB: AA 333 - 481
# ParC: AA 51 - 170
# parE: AA 366 - 523
# Source: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1142146/pdf/cjvr68pg229.pdf
fix_gyr_par_results <- function(df) {
  df <- df %>%
    mutate(gyrA_result = mut %>% # control for mutation within QRDR for gyrA
             str_extract_all("\\d+") %>% # from reprex package
             map(as.integer) %>% # converts all to integer
             map_lgl(~ any(.x >= 67L & .x <= 106L)), # returns TRUE/FALSE whether value is within range or not
           gyrA_result = if_else(gene != "gyrA", NA, gyrA_result), # filters out results for all other genes
           gyrA_result = as.integer(gyrA_result),
           gyrB_result = mut %>% # control for mutation within QRDR for gyrB
             str_extract_all("\\d+") %>%
             map(as.integer) %>%
             map_lgl(~ any(.x >= 333L & .x <= 481L)),
           gyrB_result = if_else(gene != "gyrB", NA, gyrB_result),
           gyrB_result = as.integer(gyrB_result),
           parC_result = mut %>% # control for mutation within QRDR for parC
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 51L & .x <= 170L)),
           parC_result = if_else(gene != "parC", NA, parC_result),
           parC_result = as.integer(parC_result),
           parE_result = mut %>% # control for mutation within QRDR for parE
             str_extract_all("\\d+") %>% 
             map(as.integer) %>% 
             map_lgl(~ any(.x >= 366L & .x <= 523L)),
           parE_result = if_else(gene != "parE", NA, parE_result),
           parE_result = as.integer(parE_result)) %>%
    mutate(result_gyr_par = case_when(gene == "gyrA" ~ gyrA_result,
                                      gene == "gyrB" ~ gyrB_result,
                                      gene == "parC" ~ parC_result,
                                      gene == "parE" ~ parE_result)) %>%
    mutate(result_total = ifelse(gene %in% c("gyrA","gyrB","parC","parE"), result_gyr_par, result)) %>%
    select(-c(gyrA_result,
              gyrB_result,
              parC_result,
              parE_result,
              result_gyr_par,
              result))
  return(df)
}

# Function that handles resfinder data and returns a data frame with 
# presence/absence for acquired genes                           
create_acquired_table <- function(df) {
  df <- df %>%
    select(ref, gene_names, flag, ref_ctg_change) %>%
    mutate(id = 1:n()) %>%
    filter(flag %in% flag_selection) %>%
    spread(gene_names, ref_ctg_change) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate(ref = gsub("^(.*?)_amr_report.tsv$", "\\1", ref)) %>%
    select(-c(id, flag)) %>%
    gather(gene, result,-ref) %>%
    mutate(result_total = ifelse(result == "", 0, 1),
           result_total = as.character(result_total),
           type = "gene") %>%
    select(-result)
  return(df)
}

# Function that returns a filtered dataframe based on the string "genes"
filter_mut_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in mut_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = as.character(gene))
  return(df)
}

# Function that returns a filtered dataframe based on the string "acquired_genes"
filter_acquired_table <- function(df) {
  report_genes <- unique(df$gene)
  grep_genes <- c()
  for (gene in ac_genes) {
    for (g in report_genes) {
      if (grepl(gene, g, ignore.case = T) == TRUE) {
        grep_genes <- c(grep_genes, g)
      }
    }
  }
  df <- df %>%
    filter(gene %in% grep_genes) %>%
    mutate(gene = gsub("_", "", gene),
           gene = as.character(gene))
  return(df)
}

# calculates percentage of present/absent mutations and genes
calc_stats <- function(df) {
  df <- df %>%
    group_by(gene, result_total) %>%
    count() %>%
    ungroup() %>%
    mutate(result_total = if_else(result_total == 1, "Present", "Absent")) %>%
    spread(result_total, n, fill = 0) %>%
    rowwise() %>%
    mutate(Total = Present + Absent,
           Percent = round(Present/Total*100, 1),
           lwr = round(get_binCI(Present, Total)[1], 1),
           upr = round(get_binCI(Present, Total)[2], 1))
  return(df)
}

# calculates how many mutations are present in the genes
# gyrA, gyrB, parC and parE
calc_no_of_mut <- function(df) {
  df1 <- df %>%
    filter(gene %in% c("gyrA","parC","gyrB","parE")) %>%
    mutate(entries = sapply(strsplit(.$mut, ","), FUN = function(x) {length(x)})) %>%
    separate(mut, into = as.character(c(1:max(.$entries)))) %>%
    select(-entries) %>%
    gather(id, mut, -c(ref, gene, result, type))

  df2 <- fix_gyr_par_results(df1)

  df3 <- df2 %>%
    mutate(test = ifelse(mut != "0" & result_total == 0, 0, 1)) %>%
    filter(test == 1) %>%
    spread(gene, mut) %>%
    group_by(ref) %>%
    summarise_all(funs(func_paste)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "0", "", x)) %>%
    mutate(mut_gyrA = sapply(strsplit(.$gyrA, ","), FUN = function(x) {length(x)}),
           mut_gyrB = sapply(strsplit(.$gyrB, ","), FUN = function(x) {length(x)}),
           mut_parC = sapply(strsplit(.$parC, ","), FUN = function(x) {length(x)}),
           mut_parE = sapply(strsplit(.$parE, ","), FUN = function(x) {length(x)})) %>%
    select(-c(type, id, result_total, test)) %>%
    mutate_at(.vars = vars(c("gyrA","gyrB","parC","parE")),
              .funs = function(x) ifelse(x == "", "0", x))
  return(df3)
}

# creates a data frame with one row per sample, and 1/0 results 
# for mutations in respective genes in columns
create_mut_report <- function(df) {
  df <- df %>%
    select(-mut) %>%
    group_by(ref) %>%
    mutate(id = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, type))
  return(df)
}

# creates a data frame with one row per sample, and 1/0 results
# for acquired genes in columns
create_acquired_report <- function(df) {
  df <- df %>%
    group_by(ref) %>%
    mutate(id = 1:n()) %>%
    spread(gene, result_total) %>%
    summarise_all(funs(func_paste)) %>%
    select(-c(id, type))
  return(df)
}

# prepares data frame for number of mutations heatmap
create_heatmap_df <- function(df) {
  qnr <- names(df)

  qnr_cols <- c()

  for (i in qnr) {
    if (grepl("qnr", i, ignore.case = TRUE) == TRUE) {
      qnr_cols <- c(qnr_cols, i)
    }
  }

  qnr_df <- df %>%
    select_at(vars(ref, qnr_cols)) %>%
    gather(gene, result, -ref) %>%
    mutate(type = "Gene")

  qnr_mut_quant <- mut_quant %>%
    select(-c("gyrA","gyrB","parC","parE")) %>%
    mutate(type = "Mut") %>%
    gather(gene, result, -c(ref, type)) %>%
    mutate(gene = case_when(gene == "mut_gyrA" ~ "gyrA",
                            gene == "mut_gyrB" ~ "gyrB",
                            gene == "mut_parC" ~ "parC",
                            gene == "mut_parE" ~ "parE",
                            TRUE ~ gene)) %>%
    rbind(., qnr_df)
  return(qnr_mut_quant)
}

# -------------------------------------- Analysis
# Import data
mut_data <- get_ariba_data(megares_report_loc)
acquired_data <- get_ariba_data(resfinder_report_loc)

# Clean data
clean_mut_data <- fix_gene_names(mut_data)
clean_acquired_data <- fix_gene_names(acquired_data)

# Check flags
mut_flags <- check_flags(clean_mut_data)
acquired_flags <- check_flags(clean_acquired_data)

# Wrangle
mut_table <- create_mut_table(clean_mut_data)
mut_table_fixed <- fix_gyr_par_results(mut_table)
mut_filtered <- filter_mut_table(mut_table_fixed)
acquired_table <- create_acquired_table(clean_acquired_data)
acquired_filtered <- filter_acquired_table(acquired_table)

# Stats
if (length(mut_genes) == 1) {
  if (mut_genes == "ALL") {
    mut_stats <- calc_stats(mut_table_fixed)
    mut_report <- create_mut_report(mut_table_fixed)
    }
} else {
  mut_stats <- calc_stats(mut_filtered)
  mut_report <- create_mut_report(mut_filtered)
}

if (length(ac_genes) == 1) {
  if (ac_genes == "ALL") {
    acquired_stats <- calc_stats(acquired_table)
    acquired_report <- create_acquired_report(acquired_table)
  }
} else {
  acquired_stats <- calc_stats(acquired_filtered)
  acquired_report <- create_acquired_report(acquired_filtered)
}

mut_quant <- suppressWarnings(calc_no_of_mut(mut_table))
heatmap_df <- create_heatmap_df(acquired_report)

# ---------------------------------- Tables

# number of mutations table
mut_no <- mut_quant %>%
  group_by(mut_gyrA, mut_gyrB, mut_parC, mut_parE) %>%
  count()

# number of isolates per mutation combination
mut_comb <- mut_quant %>%
  group_by(gyrA, gyrB, parC, parE) %>%
  count()
```

# Tables
## Isolate data
```{r}
isolate_data_complete <- isolate_data %>%
  left_join(no_of_reads, by = "id")

isolate_data_complete
```

## List of genes of interest
```{r}

acquired_ref_names <- acquired_data$ref_name
acquired_ref <- c()

for (i in ac_genes) {
  for (j in acquired_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      acquired_ref <- c(acquired_ref, j)
    }
  }
}

acquired_ref <- unique(acquired_ref)

acquired_ref_genes <- acquired_data %>%
  mutate(test = duplicated(ref_name),
         gene = case_when(grepl("qnra1", ref_name, ignore.case = T) == TRUE ~ "qnrA1",
                          grepl("qnrb19", ref_name, ignore.case = T) == TRUE ~ "qnrB19",
                          grepl("qnrb6", ref_name, ignore.case = T) == TRUE ~ "qnrB6",
                          grepl("qnrb60", ref_name, ignore.case = T) == TRUE ~ "qnrB60",
                          grepl("qnrs1", ref_name, ignore.case = T) == TRUE ~ "qnrS1",
                          grepl("qnrs2", ref_name, ignore.case = T) == TRUE ~ "qnrS2",
                          grepl("qnrs4", ref_name, ignore.case = T) == TRUE ~ "qnrs4",
                          grepl("qnrs9", ref_name, ignore.case = T) == TRUE ~ "qnrS9")) %>%
  filter(test == FALSE,
         ref_name %in% acquired_ref_names) %>%
  select(ref_name, gene) %>%
  mutate(compl = complete.cases(.)) %>%
  filter(compl == TRUE) %>%
  mutate(ref_name2 = sub(".+_(.*?)", "\\1", ref_name)) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = "Quinolone resistance protein, plasmid mediated, decreases susceptibility of quinolones",
         Type = "Acquired")

df_ref_names <- mut_data$cluster
ref_names <- c()

for (i in mut_genes) {
  for (j in df_ref_names) {
    if (grepl(i,j, ignore.case = T) == TRUE) {
      ref_names <- c(ref_names, j)
    }
  }
}

ref_names <- unique(ref_names)

intrinsic_ref_genes <- mut_data %>%
  mutate(test = duplicated(cluster),
         gene = case_when(grepl("gyra", cluster, ignore.case = T) == TRUE ~ "gyrA",
                          grepl("gyrb", cluster, ignore.case = T) == TRUE ~ "gyrB",
                          grepl("parc", cluster, ignore.case = T) == TRUE ~ "parC",
                          grepl("pare", cluster, ignore.case = T) == TRUE ~ "parE",
                          grepl("soxs", cluster, ignore.case = T) == TRUE ~ "soxS",
                          grepl("marR", cluster, ignore.case = T) == TRUE ~ "marR",
                          grepl("tolc", cluster, ignore.case = T) == TRUE ~ "tolC",
                          grepl("acrr", ref_name, ignore.case = T) == TRUE ~ "acrR")) %>%
  filter(test == FALSE,
         cluster %in% ref_names) %>%
  select(ref_name, gene, free_text) %>%
  mutate(ref_name2 = sub("^.*?_([A-Z]+[0-9_.]*[0-9]).*", "\\1", ref_name)) %>%
  select(-ref_name) %>%
  group_by(gene) %>%
  summarise_all(funs(func_paste)) %>%
  select(gene, ref_name2) %>%
  rename("Accession number(s)" = ref_name2,
         "Gene" = gene) %>%
  mutate(Description = case_when(Gene == "acrR" ~ "Multidrug efflux regulator, represses the transcription of acrB.",
                                 Gene == "gyrA" ~ "DNA gyrase subunit A, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "gyrB" ~ "DNA gyrase subunit B, regulates DNA topoplogy. Target for quinolones.",
                                 Gene == "marR" ~ "Repressor of the marRAB operon, involved with the activation of resistance- and oxidative stress genes.",
                                 Gene == "parC" ~ "Topoisomerase IV subunit C, regulates DNA topology. Target for quinolones.",
                                 Gene == "parE" ~ "Topoisomerase IV subunit E, regulates DNA topology. Target for quinolones.",
                                 Gene == "soxS" ~ "Transcriptional activator of the superoxide response regulon.",
                                 Gene == "tolC" ~ "Part of the major AcrAB-TolC multidrug efflux pump."),
         Type = "Intrinsic gene")

ref_genes <- rbind(intrinsic_ref_genes, acquired_ref_genes)

ref_genes

```

## Isolates per species
```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  count()

per_species
```

# Figures
## Percent QREC isolates with mutations in intrinsic genes
```{r}
gene_names <- names(mut_report)

gene_names <- gene_names[-1]

mut_report %>%
    mutate(
      ref = gsub("(.*?)\\_.+", "\\1", ref),
      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
    ) %>%
    arrange(desc(ref)) %>%
    mutate(id = id) %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_x_discrete(limits = c(
      "gyrA", "gyrB", "parC", "parE",
      "marR", "soxS", "tolC"
    )) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

## Percent QREC isolates with Qnr genes per species
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")

acquired_report %>%
    mutate(
      ref = gsub("(.*?)\\_.+", "\\1", ref),
      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
    ) %>%
    arrange(desc(ref)) %>%
    mutate(id = id) %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, qnr_genes) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(0.9)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(0.9)
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates",
         fill = NULL) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.justification = c(0, 1),
      legend.position = c(0.82, 0.97)
    )
```